---
title: "Task"
output: pdf_document
date: "`r Sys.Date()`"
---

## Packages

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(DescTools)
library(gmodels)
library(ggplot2)
library(GGally)
library(gridExtra)
library(tidyr)
library(patchwork)
library(bestNormalize)
library(caret)
library(pROC)
```

## Pre-defined functions

```{r}
create_bar_plot <- function(data, var_name) {
  freq_table <- data %>%
    group_by(!!sym(var_name), subscribed) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(!!sym(var_name)) %>%
    mutate(prop = count / sum(count))
  
  p <- ggplot(freq_table, aes(x = !!sym(var_name), y = prop, fill = subscribed)) +
    geom_bar(stat = "identity", position = "fill") +
    scale_fill_manual(values = c("TRUE" = "#4CAF50", "FALSE" = "#F44336")) +
    labs(x = var_name,
         y = "",
         fill = "Subscribed") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(size = 10),
          legend.position = "none")
  
  return(p)
}
```


## Importing data

```{r message=FALSE, warning=FALSE}
bank_full <- read_delim("bankData/bank-full.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
```

## Data cleaning

```{r}
head(bank_full)
```

```{r}
tail(bank_full)
```

```{r}
str(bank_full)
```

A first look at the data shows us that many of the provided columns have an incorrect data type. For example, default and marital status are set as character data types when they should be factors.

```{r}
bank_full <- bank_full %>%
  mutate(job = if_else(job == "admin.", "admin", job)) %>% 
  mutate(across(c(marital, job, education, contact, poutcome, day), as.factor)) %>% 
  mutate(across(c(default, housing, loan, y), ~ .x == "yes")) %>% 
  mutate(month = factor(month, levels = c("jan", "feb", "mar", "apr", "may", "jun",
                                          "jul", "aug", "sep", "oct", "nov", "dec")),

         job = relevel(job, ref = "unemployed"),
         marital = relevel(marital, ref = "single"),
         education = relevel(education, ref = "unknown"),
         contact = relevel(contact, ref = "unknown"),
         poutcome = relevel(poutcome, ref = "unknown"))

str(bank_full)
```

We have 45211 rows and 16 columns (excluding y).

A look at the description by the researchers tells us that there are no missing values even though some columns have values "unknown". We have to decide whether to keep them as "unknown" or convert them to NA. Either way, missing values must be inspected.

```{r}
sum(apply(bank_full == "unknown", 1, any))
```

There are a total of 37369 rows with at least one "unknown" value.

How many "unknowns" does each column have?

```{r}
unknown_table <- data.frame(
  unknown_count = sapply(bank_full, function(col) sum(col == "unknown", na.rm = TRUE))) %>% 
  arrange(desc(unknown_count)) %>% 
  filter(unknown_count != 0)

print(unknown_table)
```

Almost all of the poutcome values are unknown. Let's keep this column for now as we will look at outcome distributions with regard to y values later on.

Lastly, since some columns have names that may be difficult to interpret without looking at the metadata first, we should rename them.

```{r}
bank_full <- bank_full %>% 
  rename(in_default = "default",
         housing_loan = "housing",
         personal_loan = "loan",
         contact_type = "contact",
         subscribed = "y")
```

```{r}
lapply(bank_full[ , !(names(bank_full) %in% c("age", "balance", "duration", "pdays"))], unique)
```
Looking at the unique columns values we do not see anything out of the ordinary.

## Exploratory analysis

Now we can look at each variable separately.

### Subscribed

```{r}
table(bank_full$subscribed)
```

We have a large imbalance in out data. Only 11,6% of contacted clients subscribed. We must also take this into account when removing unknown values.

### Age

```{r}
ggplot(bank_full, aes(x = age)) +
  geom_bar() + 
  facet_grid(subscribed ~ ., scales = "free_y")
```
The vast majority of clients contacted by the bank were between 26 and 60 years old. The variable is not follow the normal distribution.

```{r}
bank_full = bank_full %>% 
  mutate(age_categ = case_when(
    age > 60 ~ "high",
    age > 26 ~ "mid",
    TRUE ~ "low"
  ))

CrossTable(bank_full$subscribed, bank_full$age_categ, prop.t = FALSE, prop.chisq = FALSE)
```
Clients of the age 60+ were most likely to subscribe: 42.3% of them chose to do so. That is the highest percentage of all age groups even though older clients make up the smallest part of the total population.

The data does not seem to indicate a linear relationship between age and subscription rates. Either way, we will keep a continuous version of the age variable.

```{r}
vars_for_plot <- bank_full %>%
  select(age, balance, duration, campaign, pdays, previous, subscribed)

bank_long <- vars_for_plot %>%
  pivot_longer(cols = -subscribed, names_to = "variable", values_to = "value")

ggplot(filter(bank_long, variable == "age"),
                aes(x = value, fill = subscribed)) +
  geom_density(alpha = 0.5) +
  xlim(18, 99)
```
The density plots also do not show a large difference in terms of age with the acception being clients over the age of 60.

### Job

```{r}
summary(bank_full$job)
```
There are a total of 228 unknown values. Due to the large row count we can afford to drop these rows.

```{r}
bank_full <- bank_full %>% filter(job != "unknown")
```

```{r}
nrow(bank_full)
```

Let's look at what percentage of clients subscribed based on their job.

```{r}
create_bar_plot(bank_full, "job")
```

As the chart shows, students, of all jobs, were most likely to subscribe to a deposit with retired workers following second.

### Marital status

```{r}
CrossTable(bank_full$subscribed, bank_full$marital, prop.t = FALSE, prop.chisq = FALSE)

```

Married persons make up 60.1% of out data set. Single clients were slightly more likely to make a subscription (14,9%) than other clients. It is also probable that this tendency is caused by randomness as marital status categories are not divided equally.

```{r}
create_bar_plot(bank_full, "marital")
```

### Education

```{r}
CrossTable(bank_full$subscribed, bank_full$education, prop.t = FALSE, prop.chisq = FALSE)
```

There are "unknown" values in the education variable. If we removed these "unknowns" we would risk causing further imbalance in the subscribed variable as only around 20% of clients decided to make a deposit subscription. It is most probable that clients with a tertiary education will subscribe.

```{r}
create_bar_plot(bank_full, "education")
```


### Default status

```{r}
CrossTable(bank_full$subscribed, bank_full$in_default, prop.t = FALSE, prop.chisq = FALSE)
```

Only 6,4% of clients that were in default chose to make a subscription. Out of the total sample only 1,8% clients were in default. This variable is unlike to be a good indicator of wether the client makes a subscription.

```{r}
create_bar_plot(bank_full, "in_default")
```

### Balance

```{r message=FALSE, warning=FALSE}
ggplot(filter(bank_long, variable == "balance"),
                    aes(x = value, fill = subscribed)) +
  geom_density(alpha = 0.5) +
  xlim(-1000, 5500)
```

The distribution of balance does not immediately indicate that wealthier clients are more likely to make a subscription.

```{r message=FALSE, warning=FALSE}
ggplot(filter(bank_long, variable == "balance"),
                    aes(x = value, fill = subscribed)) +
  geom_boxplot(alpha = 0.5) +
  xlim(-1000, 5500)
```
Since we are dealing with financial data the are many exceptions (outliers) in the distributions of variables. Though the box plots do indicate that the median of balance is higher for those who chose to subscribe.

```{r}
paste0("Balance Mean: ", mean(bank_full$balance, na.rm = TRUE))
paste0("Balance Standart Deviation: ", sd(bank_full$balance))

outliers <- boxplot.stats(bank_full$balance)$out
outlierNum <- length(outliers)
paste0("Outlier Percentage: ", round(outlierNum/(length(bank_full$balance)) * 100, 2))

```

Since the balance variance is relatively high (3035) and 10.49% of the entries can be marked as outliers, we'll normalize the balance variable using the Order-Norm transformation (maps each data point to a percentile in a normal distribution based on the percentile value in the original distribution).

```{r warning=FALSE}
on <- orderNorm(bank_full$balance)
bank_full$trans_balance <- predict(on)

ggplot(bank_full, aes(x = job, y = trans_balance, fill = job)) + 
  geom_boxplot(outlier.size = 0.7, na.rm = TRUE) +
  coord_flip() + 
  stat_summary(fun = mean, geom = "point", shape = 4, size = 0.8, color = "black", na.rm = TRUE) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```

The box plots allow us to conclude that the balance of the client accounts is likely dependent more factors than simply their job. It also indicates that the clients grouped by their job type are not homogeneous (as we had to apply Order-Norm transformation to achieve more normal values). Nevertheless we can draw certain conclusions. For example, we can see that the median account balance of students is higher than those of service workers. Another trend is clear - retirees have the highest average and median balance.

### Housing and Personal loans

```{r}
CrossTable(bank_full$subscribed, bank_full$housing_loan, prop.t = FALSE, prop.chisq = FALSE)
```

55,9% of the clients in out sample had a housing loan. Clients that did not have a housing loan were more twice as likely to subscribe than the clients without one. It is clear that this variable will be significant when modelling.

```{r}
CrossTable(bank_full$subscribed, bank_full$personal_loan, prop.t = FALSE, prop.chisq = FALSE)
```

The situation is practically the same with personal loans accept here only 16,1% of the clients had a personal loan. Clients that did not have a personal loan were twice as likely to subscribe than the clients without one. It is clear that this variable will also be significant when modelling as clients with no financial burdens (defaults and loans) are more likely to subscribe.

```{r}
plot_list <- lapply(c("housing_loan", "personal_loan"), function(var) create_bar_plot(bank_full, var))

bar_plot_matrix <- grid.arrange(grobs = plot_list, ncol = 2)
```

### Contact type

```{r}
CrossTable(bank_full$subscribed, bank_full$contact_type, prop.t = FALSE, prop.chisq = FALSE)
```

Clients that were contacted through a cellular phone were slightly more likely to make a subscription. The contact type for 28,7% of the clients is unknown.

```{r}
create_bar_plot(bank_full, "contact_type")
```

### Day and month

```{r}
plot_list_2 <- lapply(c("day", "month"), function(var) create_bar_plot(bank_full, var))

bar_plot_matrix_2 <- grid.arrange(grobs = plot_list_2, ncol = 2)
```
March, September, October and December were the best months to contact the clients. Higher success could also be achieved when contacting the clients on the 1st, 10th, 22nd and 30th. These insights should be tested when modelling.

### Duration

```{r message=FALSE, warning=FALSE}
ggplot(filter(bank_long, variable == "duration"),
                     aes(x = value, fill = subscribed)) +
  geom_density(alpha = 0.5) +
  xlim(0, 1600)
```
Call duration seems to tell a clearer story than other continuous variables. Clients that, in the end, decided not to subscribe had shorter conversations with the representative of the bank showing their disinterest early on.

### Atributes related to previous contact

#### Campaign contacts

```{r message=FALSE, warning=FALSE}
ggplot(bank_full, aes(x = campaign)) +
  geom_bar() + 
  facet_grid(subscribed ~ ., scales = "free_y") +
  xlim(0, 25)
```
Number of contacts performed during this campaign seems to be proportional with the number of contacts performed in total.

Let's took how the number of total calls is related to calls that lead to a subscription. 

```{r}
CrossTable(bank_full$subscribed[bank_full$campaign < 5], bank_full$campaign[bank_full$campaign < 5],
           prop.t = FALSE, prop.chisq = FALSE)
```
Number of contacts during the campaign seems to increase the likeliness of subscription but with linearly diminishing returns.

#### Previous days

```{r}
sum(bank_full$pdays != -1)
```
There are 8224 clients which have been contacted in the past. Since there are many different pdays values and since the variable has been encoded as -1 or any other natural number, in order to avoid singularities in our logistic regression model, we can transform this variable in to a binary variable.

```{r}
bank_full <- bank_full %>%
  mutate(was_contacted = ifelse(pdays == -1, FALSE, TRUE))
```

#### Previous contacts

```{r message=FALSE, warning=FALSE}
ggplot(bank_full, aes(x = previous)) +
  geom_bar() + 
  facet_grid(subscribed ~ ., scales = "free_y") + 
  xlim(-1, 50)
```

```{r}
CrossTable(bank_full$subscribed[bank_full$previous < 4], bank_full$previous[bank_full$previous < 4],
           prop.t = FALSE, prop.chisq = FALSE)
```
Number of contacts during the previous campaign seems to linearly increase the likeliness of subscription.

#### Previous outcome

```{r}
CrossTable(bank_full$subscribed, bank_full$poutcome, prop.t = FALSE, prop.chisq = FALSE)
```

If the outcome of the previous campaign was successful, the outcome of the current campaign on the same client has a 64,6% likelihood of being successful. Although it must be noted that there are only 1500 clients with the poutcome attribute set as successful.

```{r}
create_bar_plot(bank_full, "poutcome")
```

### Corelation of continuous variables

```{r}
corr_matrix <- cor(bank_full[, c("age", "balance", "duration")], use = "complete.obs")
print(corr_matrix)
```
As the continuous variables are not correlated with each other, we can negate multicollinearity concerns for the logistic regression model.

## Manipulating data

We select a small random sample of the provided data with a pre-determined seed for repeatable results.

```{r}
set.seed(167)
smallBank <- sample_n(bank_full, 400, replace = FALSE)
```

Let's choose a data frame with the clients that have a dangerously low balance and have or have had a partner at a point in their life. Due to low numbers in the total population, let's search for them in the full data set.

```{r}
lowBalwPartner <- bank_full %>% 
  filter(balance < 100 & marital %in% c("maried", "divorced"))
```

Also, we'll filter another group of clients which have at least one loan with the bank and are at least of the median age for the data set.

```{r}
withLoans <- bank_full %>% 
  filter((housing_loan == TRUE | personal_loan == TRUE) & age >= median(age, na.rm = TRUE))
```

We may also calculate the summarizing statistics.

```{r}
job_summary <- bank_full %>%
  group_by(job) %>%
  summarise(
    age_mean = round(mean(age, na.rm = TRUE), 2),
    balance_mean = mean(balance, na.rm = TRUE),
    balance_median = median(balance, na.rm = TRUE),
    balance_sd = sd(balance, na.rm = TRUE),
    duration_median = median(duration, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(n), desc(age_mean))

print(job_summary)
```

The summarized statistics allows us to make a few insights about the clients that were contacted. First, the clients with a job in management had the highest average balance. Second, high standard deviation tells us that client balance varies quite a lot from one client to another. Third, most clients over all had a balance in the mid-500s. Fourth, most of the contacted clients were blue-collar workers. That is quite normal as blue-collar workers usually make up the largest percentage of the population.

We should also inspect the clients that chose to subscribe to a deposit and what characteristics they show.

```{r}
subscriber_summary <- bank_full %>%
  filter(subscribed == TRUE) %>%
  select(-in_default) %>%
  summarise(across(everything(), ~DescTools::Mode(.x), .names = "mode_{.col}"))

print(subscriber_summary)
```

The data shows us that the "most common" client that chose to subscribe to a deposit is a 32 y.o. married management worker which was contacted via phone in May and the phone call lasted 261 seconds. These could be the key factors which influence the probability of subscription.

Using the previous conclusion, we may create a mock variable that assigns a score of how likely each client is to subscribe to a deposit. In order to give sense to the number representation of the score, we will apply a min-max transformation.

```{r}
find_engagement <- function(duration, balance, housing_loan, personal_loan, in_default) {
  if(in_default != TRUE){
    score <- duration + 10 * (balance / 1000) - housing_loan * 10 - personal_loan * 20
    if (score < 0){
      return(0)
    } else {
      return(score)
    }
  } else {
    return(0)
  }
}

bank_full <- bank_full %>% 
  mutate(engagement_score = mapply(find_engagement, duration, balance, housing_loan, personal_loan, in_default)) %>%
  mutate(engagement_score = round((engagement_score - min(engagement_score, na.rm = TRUE)) / 
           (max(engagement_score, na.rm = TRUE) - min(engagement_score, na.rm = TRUE)), 3))
```

In order to detect clients that have no loans and sufficient balance to make a bank term deposit (a. k. a. are "good" potential depositors), but have specifically chosen not to, we will create a new indicator column.

```{r}
bank_full_potencial <- bank_full %>%
  mutate(potential_client = balance > 1000 & campaign > 0 & previous == 0 &
                            !in_default & !housing_loan & !personal_loan)
summary(bank_full_potencial$potential_client)
```

We can see that to 5227 "potential" clients the marketing campaign hasn't been effective.


## Modelling

Next, we have to create dummy variables for categorical columns.

```{r}
dmy <- dummyVars(~ age + age_categ + job + marital + education + balance + contact_type + day + month + campaign + pdays + previous + poutcome + duration, data = bank_full[ , names(bank_full) != "subscribed"])

dummy_data <- data.frame(predict(dmy, newdata = bank_full))

dummy_data <- dummy_data[, setdiff(colnames(dummy_data), c("contact_type.unknown", "job.unemployed", "month.jan", "marital.single", "education.unknown", "poutcome.unknown", "day.1"))]

dummy_full <- cbind(dummy_data, subscribed = bank_full$subscribed, in_default = bank_full$in_default, housing_loan = bank_full$housing_loan, personal_loan = bank_full$personal_loan)
```

We can now separate our original data set into two: training and testing.

```{r}
set.seed(167)
sample_size <- round(0.8 * nrow(dummy_full))

train_indices <- sample(seq_len(nrow(dummy_full)), size = sample_size)

train_dummy <- dummy_full[train_indices, ]
test_dummy  <- dummy_full[-train_indices, ]
```

And finally, we can run the model.

```{r}
model1 <- glm(subscribed ~ ., data = train_dummy, family = binomial)
summary(model1)
```

```{r}
glm_predict_subs <- predict(model1, test_dummy, type = "response")

roc_curve <- roc(test_dummy$subscribed, glm_predict_subs)

auc(roc_curve)
```

