---
title: "Data Manipulation Task"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
```

## Importing data

```{r}
bank_full <- read_delim("bankData/bank-full.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
```

## Data cleaning

```{r}
head(bank_full)
```

A first look at the data shows us that many of the provided columns have an incorrect data type. For example, default and marital status are set as character data types while they should be factors.

```{r}
str(bank_full)
```

A look at the description by the researchers tells us that there are no missing values even though some columns have values "unknown". That also should be changed.

```{r}
bank_full[bank_full == "unknown"] <- NA

bank_full <- bank_full %>%
  mutate(job = if_else(job == "admin.", "admin", job)) %>% 
  mutate(across(c(marital, job, education, contact), as.factor)) %>% 
  mutate(across(c(default, housing, loan), ~ .x == "yes"))

str(bank_full)
```
